The effect of linking assumptions and number of response options on inferred scalar implicature rate

### 2018-05-08 writeup.Rmd
rewrote quantitative results section

Our primary goal in this study was to check whether the estimated "implicature rate" in the experimental task is affected by the linking assumption and the number of response options available in the task. Our analysis in this section focuses on these three elements. As mentioned before, two trial types were predicted to include pragmatic implicatures. First, trials where two animals were on the card but the fictional character guessed with a disjunction (scalar); for example "cat or dog" when the card has both a cat and a dog on it. Second, trials where there were two animals on the card but the character guessed only one (exhaustive); for example "cat" when the card had a cat and a dog on it. We called such trials "exhaustive". In our assessment of implicature rate, we focus on these two trial types.

We considered two linking assumptions. First we defined a weak (lenient) linking assumption in which any response lower than the maximum point on the scale (i.e. "right") is considered evidence for implicature computation. Second, we defined a strong (strict) linking assumption that only considered the lowest point on the scale (i.e. "wrong") as evidence for implicature computation. For each condition in our study (binary, ternary, quaternary, and quinary) and each implicature trial type (exhaustive and scalar), we computed a weak and a strong implicature rate. Figure \@ref(fig:implicatureRatePlot) shows these computed rates.

Comparing the strong and weak rows on Figure \@ref(fig:implicatureRatePlot), we see that a weak linking assumption tends to estimate higher implicature rates, especially in tasks with more response options. With a strong linking assumption, the binary and possibly ternary judgment tasks derive higher implicature rates than quaternary and quinary tasks. With a weak linking assumption, the pattern is reversed. Quaternary and quinary tasks estimate higher rates than binary and ternary tasks. The patterns show that estimates of "implicature rate" depend on linking assumptions and the number of responses available to participants in the study.

Comparing the exhaustive and scalar columns of Figure \@ref(fig:implicatureRatePlot), we see that with a strong linking assumption, there are slightly higher rates for scalar implicatures in the binary and ternary tasks. With a weak linking assumption, there may be slightly higher rates for scalar implicatures in the binary and ternary while the rates may be lower in the quaternary and quinary tasks. In what follows, we formally test the effect of linking assumption and response options on exhaustive and scalar implicature rates.

```{r implicatureRatePlot, fig.width=5, fig.height=3, fig.cap = "Inferred implicature rate on exhaustive and scalar trials with the binary, ternary, quaternary, and quinary response task. Columns indicate link from response to implicature rate (strong: proportion of "wrong" judgments; weak: proportion of non-"right" judgments)."}
ggplot(implicature_results, aes(x=response_type,y=proportion)) +
  geom_bar(aes(fill=definition), stat="identity") +
  facet_grid(trial_type~definition, drop = TRUE) +
  theme_few() + 
  labs(x="Response task", y="Inferred implicature rate") +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_linerange(aes(ymin=lower,ymax=upper)) + guides(fill=FALSE)
```

@@

fixed_effects_table <- 
  fixed_effects %>%
  rename(Predictors = "X1", `2.5%` = "X2.5.ile", `97.5%` = "X97.5.ile") %>%
  select(-Est.Error)

kable(fixed_effects_table,digits = 2 , caption = "\\label{tab:modeltable}Model Parameter Estimates and Their Credible Intervals", booktabs = T)
```

For our formal analysis, we used a bayesian binomial mixed effects model using the R packge "brms" [@burkner2016brms]. The model had the fixed effects of response type (binary, ternary, quaternary, quinary), linking assumption (strong vs. weak), and trial type (exhaustive vs. scalar), as well as their two way and three way interactions. We also included random intercepts and slopes for items (cards) and participants. Following @barr2013random, we used the maximal random effects structure by including random slopes for all our fixed effects and their interactions. Since the number of response options was a between participant variable we did not include random slopes of response options for participants. Four chains converged after 2000 iterations each (warmup = 1000). Table \@ref(tab:modeltable) summarizes the mean parameter estimates and their 95% credible intervals. $\hat{R}=1$ for all estimated parameters. All the analytical decisions described here were pre-registered[^3].

The model provided evidence for two effects in the study. First, there was a main effect of trials such that scalar trials had slightly higher implicature rates than exhaustive trials (Mean Estimate = `r round(fixed_effects$Estimate[6],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[6],2)`, `r round(fixed_effects$X97.5.ile[6],2)`]). Second, there was an interaction between linking assumption and number of resopnse options such that the quaternary task (Mean Estimate = `r round(fixed_effects$Estimate[7],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[7],2)`, `r round(fixed_effects$X97.5.ile[7],2)`]) and the quinary task  (Mean Estimate = `r round(fixed_effects$Estimate[8],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[8],2)`, `r round(fixed_effects$X97.5.ile[8],2)`]) with a weak linking assumption had higher implicature rates.

>>>>

Our primary goal in this study was to test whether the estimated implicature rate in the experimental task is affected by the linking assumption and the number of response options available to participants. To this end, we only analyzed the critical trials (exhaustive and scalar). In particular, we classified each data point from critical trials as constituting an implicature (1) or not (0) under the strong and weak link. Figure \@ref(fig:implicatureRatePlot) shows the resulting implicature rates by condition and link.

Visually, the weak link tends to result in greater estimates of implicature rates, especially in tasks with more response options. Under the strong link, this latter pattern is reversed: the binary and ternary judgment tasks result in greater estimates of implicature rates than with more response options.

```{r implicatureRatePlot, fig.width=4, fig.height=3, fig.cap = "Inferred implicature rates on exhaustive and scalar trials as obtained with the binary, ternary, quaternary, and quinary response task. Columns indicate link from response to implicature rate (strong: proportion of 'wrong judgments; weak: proportion of non-'right' judgments)."}
ggplot(implicature_results, aes(x=response_type,y=proportion)) +
  geom_bar(aes(fill=definition), stat="identity") +
  facet_grid(trial_type~definition, drop = TRUE) +
  theme_few() + 
  labs(x="Response task", y="Inferred implicature rate") +
  theme(text = element_text(size=12), axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_linerange(aes(ymin=lower,ymax=upper)) + guides(fill=FALSE)
```

@@

fixed_effects_table <- 
  fixed_effects %>%
  rename(Predictors = "X1", `2.5%` = "X2.5.ile", `97.5%` = "X97.5.ile") %>%
  select(-Est.Error) %>%
  mutate(Evidence=sign(`2.5%`)==sign(`97.5%`)) %>%
  mutate(Evidence=replace(Evidence,Evidence==TRUE,"*")) %>%
  mutate(Evidence=replace(Evidence,Evidence==FALSE,""))

kable(fixed_effects_table,digits = 2 , caption = "\\label{tab:modeltable}Model parameter estimates and their credible intervals. Rows marked with an asterisk in the evidence column do not contain 0 in the credible interval, thereby providing evidence for an effect.", booktabs = T)
```

To analyze the effect of link and response options on inferred implicature rate, we used a Bayesian binomial mixed effects model using the R packge "brms" [@burkner2016brms] with uninformative priors **JD: Masoud is the uninformative prior part correct?**. The model predicted the log odds of implicature over no implicature from fixed effects of *response type* (binary, ternary, quaternary, quinary -- dummy-coded with binary as reference level), *link* (strong vs. weak -- dummy-coded with strong as reference level), and trial type (exhaustive vs. scalar -- dummy-coded, with exhaustive as reference level), as well as their two-way and three-way interactions. Following @barr2013random, we included the maximal random effects structure justified by the design: random intercepts for items (cards) and participants, random by-participant slopes for link, trial type, and their interaction, and random by-item slopes for link, trial type, response type, and their interactions. Since the number of response options was a between participant variable we did not include random slopes of response options for participants. Four chains converged after 2000 iterations each (warmup = 1000). Table \@ref(tab:modeltable) summarizes the mean parameter estimates and their 95% credible intervals. $\hat{R}=1$ for all estimated parameters. All the analytical decisions described here were pre-registered[^3].

The model provided evidence for the following effects: First, there was a main effect of trial type such that scalar trials resulted in greater implicature rates than exhaustive trials (Mean Estimate = `r round(fixed_effects$Estimate[6],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[6],2)`, `r round(fixed_effects$X97.5.ile[6],2)`]). Second, there was an interaction between link and number of response options such that the quaternary task (Mean Estimate = `r round(fixed_effects$Estimate[7],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[7],2)`, `r round(fixed_effects$X97.5.ile[7],2)`]) and the quinary task  (Mean Estimate = `r round(fixed_effects$Estimate[8],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[8],2)`, `r round(fixed_effects$X97.5.ile[8],2)`]) with a weak link resulted in greater implicature rates. Finally, there was a three-way interaction between link, trial type, and number of response options (Mean Estimate = `r round(fixed_effects$Estimate[15],2)`, 95% Credible Interval=[`r round(fixed_effects$X2.5.ile[15],2)`, `r round(fixed_effects$X97.5.ile[15],2)`]). One interpretation of this interaction is that the difference between the weak and strong link on scalar trials in the quinary task was smaller than on exhaustive trials, though we believe this is not too interesting, given that the binary reference level implicature estimate was lower for exhaustive trials in the first place. **JD: happy to throw out this last sentence** Crucially, both number of response options and link affect the inferred implicature rate.
